---
title: "Assignment1"
author: "Harlee Scordo"
date: "2023-10-18"
output: html_document
---

## Exercise 1:

```{r}

favorite_movie <- "Star_Wars"

x <- 6

v <- c(2, 4, 6)

v * 3

y <- 7

z<- x * y

```

## Exercise 2:

*It says [1] 9 underneath the code and in the console, but not in the environment.*

```{r}

6+3

```
## Exercise 3:

*The answer appears in the environment now.*

*When I add a, it adds the text down at the bottom again.*

```{r}

m <- 6

n <- 3

a <- m + n

a
```
## Exercise 4:

*It calculates 81. It took the previous value of 9 and adapted it here. *

```{r}
a ^2


```

## Exercise 5:

*It takes the a, x and y values (the orgional ones) and adds them altogether.* 

```{r}
sum (a, x, y)

```

## Exercise 6:

*This got rid of everything standing in the environment pane. *


## Exercise 7:

*It repopulated everything that was there before. It seems to keep it in alphabetical order. *

## Exercise 8:

*This added the total to the bottom. It also added 2 to each value for v.* 

```{r}

v +2

```


## Exercise 9:

*It states chr instead of num. *

```{r}

music <- c("bluegrass", "funk", "folk")

```

